Full Waveform Inversion by Source Extension: Why it works

### 2020-04-18 repro/papers/dso1d/paper.tex
more

Ignoring amplitude, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many descriptions of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above, which
amounts essentially to a time shift. To make the link with wavelet
frequency content manifest, I introduce a family $\{\fl\}$ of
wavelets indexed by $\lambda$, a parameter having dimensions of time,
\begin{equation}
  \label{eqn:deffl}
  \fl(t) = \frac{1}{\lambda}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
The argument $s$ of the ``mother wavelet'' $f_1$ is
nondimensional. The only constraints placed on $f_1$ are that (i) 
$f_1(s)=0$ for $|s|\ge 1$, and (ii) $f_1$  has positive
mean-square, that is, does not vanish identically. Note that
the scaling is such that the mean-square
\[
  \|\fl\|^2 = \int\,dt\,|\fl(t)|^2
\]
is independent of $\lambda$.

@@

This constraint actually rules out some popular approaches to FWI. To
begin with, basic least-squares FWI as formulated in the third section
above can be reformulated as a quadratic form whose Hessian turns out
not to be pseudodifferential. Therefore the FWI objective is not
smooth in data and model jointly, a fact that is linked to the
cycle-skipping behaviour demonstrated above. More surprising, perhaps,
the same turns out to be true for Wavefield Reconstruction Inversion (WRI),
an extended source inversion algorithm introduced by
\cite{LeeuwenHerrmannWRI:13}, and further developed by
\cite{LeeuwenHerrmann:16}, \cite{WangYingst:SEG16}, and
\cite{Aghamiry:19}, amongst others. This approach turns out to be closely
linked to basic FWI, and can be formulated as minimization of a similar quadratic form
which once again has a non-pseudodifferential Hessian. Not
coincidentally, it also exhibits cycle-skipping behaviour. Just as for
FWI, in simple cases such as the problem studied in this paper, WRI
can be shown explicitly to have local minima far from the global
minimum, and a region of attraction for the global minimum on the
order of a wavelength in diameter, just as does FWI.

>>>>

Ignoring amplitude, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many descriptions of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above, which
amounts essentially to a time shift. To make the link with wavelet
frequency content manifest, I introduce a family $\{\fl\}$ of
wavelets indexed by $\lambda$, a parameter having dimensions of time,
\begin{equation}
  \label{eqn:deffl}
  \fl(t) = \frac{1}{\sqrt{\lambda}}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
The argument $s$ of the ``mother wavelet'' $f_1$ is
nondimensional. The only constraints placed on $f_1$ are that (i) 
$f_1(s)=0$ for $|s|\ge 1$, and (ii) $f_1$  has positive
mean-square, that is, does not vanish identically. Note that
the scaling is such that the mean-square
\[
  \|\fl\|^2 = \int\,dt\,|\fl(t)|^2
\]
is independent of $\lambda$.

@@

This constraint actually rules out some popular approaches to FWI. To
begin with, basic least-squares FWI as formulated in the third section
above can be reformulated as a quadratic form whose Hessian turns out
not to be pseudodifferential. Therefore the FWI objective is not
smooth in data and model jointly, a fact that is linked to the
cycle-skipping behaviour demonstrated above. More surprising, perhaps,
the same turns out to be true for Wavefield Reconstruction Inversion (WRI),
an extended source inversion algorithm introduced by
\cite{LeeuwenHerrmannWRI:13}, and further developed by
\cite{LeeuwenHerrmann:16}, \cite{WangYingst:SEG16}, and
\cite{Aghamiry:19}, amongst others. This approach turns out to be closely
linked to basic FWI, and can be formulated as minimization of a similar quadratic form
which once again has a non-pseudodifferential Hessian. Not
coincidentally, it also exhibits cycle-skipping behaviour. Just as for
FWI, in simple cases such as the problem studied in this paper, WRI
can be shown explicitly to have local minima far from the global
minimum, and a region of attraction for the global minimum on the
order of a wavelength in diameter, just as does FWI \cite[]{wwsorcas:20-01}.
